Follow-up Samples

### 2017-01-18 submission/manuscript_outline_20161024.Rmd
streamline - abstract part 2

**Background:** Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease. Previous research suggests a link between fecal bacterial microbiome and CRC.   The overall objective of our study was to investigate the changes in the bacterial microbiome after surgery in patients with lesion (i.e. adenoma or carcinoma).  Specifically, we wanted to identify what within the community was different within those undergoing surgical removal of lesion.  We also wanted to investigate the use of the bacterial microbiome and Fecal Immunoglobulin Test (FIT) to build models which classify individuals as having a lesion or whether samples could be classified as before or after surgery.     
**Results:** Adenoma individual's bacterial microbiome were more similar to their pre-surgery sample then those with carcinoma (P-value = `r format(change_summary['thetayc', 'Pvalue'], digits = 3)`) and this was also reflected in FIT  (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`).  There was no significant difference in any indivdiual OTU between samples before and after surgery (P-value > `r format(min(filter(all_OTU_pair_wilcox, analysis == "ALL")[, 'BH_corrected']), digits = 3)`).  A model with a total of `r length(lesion_top10_varsMDA$variable)` variables was able to classify lesion (AUC = `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) while the model to classify samples as before and after surgery had `r length(IF_top10_varsMDA$variable)` variables (AUC = `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)`).  Both models had a significant decrease in the positive probability of a lesion or being a before surgery sample (P-value = `r format(filter(all_probs_paired_wilcox, model_type == "red_lesion" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)` and `r format(filter(all_probs_paired_wilcox, model_type == "red_IF" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).  In total there were `r length(common_vars_summary$otu)` OTUs that were common to both models and the majority of these classified to commensal bacteria with the largest representation belonging to `r common_vars_summary["Otu000005", "lowest_ID"]`, `r common_vars_summary["Otu000012", "lowest_ID"]`, `r common_vars_summary["Otu000082", "lowest_ID"]`, and `r paste(gsub("_unclassified", "", common_vars_summary["Otu000006", "lowest_ID"]))`.   
**Conclusions:** Our data suggests that treatment not only significantly reduces the probability of having a colonic lesion within our models but also causes detectable changes in the bacterial microbiome.  Further surveillance of these individuals will enable us to determine whether models such as the one we present here can also be used to predict recurrence of colorectal cancer.

>>>>

**Background:** Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease. Previous research suggests a link between fecal bacterial microbiome and CRC.   The overall objective of our study was to investigate the changes in the bacterial microbiome after surgery in patients with lesion (i.e. adenoma or carcinoma).  Specifically, we wanted to identify what within the community was different within those undergoing surgical removal of lesion.  We also wanted to investigate the use of the bacterial microbiome and Fecal Immunoglobulin Test (FIT) to build models which could classify individuals as having a lesion or as before surgery.     
**Results:** Adenoma individual's bacterial microbiome were more similar to their pre-surgery sample then those with carcinoma (P-value = `r format(change_summary['thetayc', 'Pvalue'], digits = 3)`).  Their change in FIT was also significantly different then those with carcinoma (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`).  There was no significant difference in any indivdiual OTU between samples before and after surgery (P-value > `r format(min(filter(all_OTU_pair_wilcox, analysis == "ALL")[, 'BH_corrected']), digits = 3)`).  A model with a total of `r length(lesion_top10_varsMDA$variable)` variables was able to classify lesion (AUC = `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) while the model to classify samples as before surgery had `r length(IF_top10_varsMDA$variable)` variables (AUC = `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)`).  The after surgery samples had a significant decrease in the model positive probability of a lesion or before surgery sample (P-value = `r format(filter(all_probs_paired_wilcox, model_type == "red_lesion" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)` and `r format(filter(all_probs_paired_wilcox, model_type == "red_IF" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).  In total there were `r length(common_vars_summary$otu)` OTUs that were common to both models and the majority of these classified to commensal bacteria (`r common_vars_summary["Otu000005", "lowest_ID"]`, `r common_vars_summary["Otu000012", "lowest_ID"]`, `r common_vars_summary["Otu000082", "lowest_ID"]`, and `r paste(gsub("_unclassified", "", common_vars_summary["Otu000006", "lowest_ID"]))`).   
**Conclusions:** Our data suggests that treatment not only significantly reduces the probability of having a colonic lesion within our models but also causes detectable changes in the bacterial microbiome.  Further surveillance of these individuals will enable us to determine whether models such as the ones we present could be used to predict recurrence of colorectal cancer.

### 2017-01-18 submission/manuscript_outline_20161024.tex
streamline - abstract part 2

\textbf{Background:} Colorectal cancer (CRC) continues to be a worldwide
health problem with early detection being used as a key component in
mitigating deaths due to the disease. Previous research suggests a link
between fecal bacterial microbiome and CRC. The overall objective of our
study was to investigate the changes in the bacterial microbiome after
surgery in patients with lesion (i.e.~adenoma or carcinoma).
Specifically, we wanted to identify what within the community was
different within those undergoing surgical removal of lesion. We also
wanted to investigate the use of the bacterial microbiome and Fecal
Immunoglobulin Test (FIT) to build models which classify individuals as
having a lesion or whether samples could be classified as before or
after surgery.\\
\textbf{Results:} Adenoma individual's bacterial microbiome were more
similar to their pre-surgery sample then those with carcinoma (P-value =
0.00198) and this was also reflected in FIT (P-value = 2.15e-05). There
was no significant difference in any indivdiual OTU between samples
before and after surgery (P-value \textgreater{} 0.125). A model with a
total of 37 variables was able to classify lesion (AUC = 0.847 - 0.791)
while the model to classify samples as before and after surgery had 33
variables (AUC = 0.79 - 0.651). Both models had a significant decrease
in the positive probability of a lesion or being a before surgery sample
(P-value = 1.91e-11 and 6.72e-12). In total there were 14 OTUs that were
common to both models and the majority of these classified to commensal
bacteria with the largest representation belonging to Bacteroides,
Blautia, Streptococcus, and Clostridiales.\\
\textbf{Conclusions:} Our data suggests that treatment not only
significantly reduces the probability of having a colonic lesion within
our models but also causes detectable changes in the bacterial
microbiome. Further surveillance of these individuals will enable us to
determine whether models such as the one we present here can also be
used to predict recurrence of colorectal cancer.

>>>>

\textbf{Background:} Colorectal cancer (CRC) continues to be a worldwide
health problem with early detection being used as a key component in
mitigating deaths due to the disease. Previous research suggests a link
between fecal bacterial microbiome and CRC. The overall objective of our
study was to investigate the changes in the bacterial microbiome after
surgery in patients with lesion (i.e.~adenoma or carcinoma).
Specifically, we wanted to identify what within the community was
different within those undergoing surgical removal of lesion. We also
wanted to investigate the use of the bacterial microbiome and Fecal
Immunoglobulin Test (FIT) to build models which could classify
individuals as having a lesion or as before surgery.\\
\textbf{Results:} Adenoma individual's bacterial microbiome were more
similar to their pre-surgery sample then those with carcinoma (P-value =
0.00198). Their change in FIT was also significantly different then
those with carcinoma (P-value = 2.15e-05). There was no significant
difference in any indivdiual OTU between samples before and after
surgery (P-value \textgreater{} 0.125). A model with a total of 37
variables was able to classify lesion (AUC = 0.847 - 0.791) while the
model to classify samples as before surgery had 33 variables (AUC = 0.79
- 0.651). The after surgery samples had a significant decrease in the
model positive probability of a lesion or before surgery sample (P-value
= 1.91e-11 and 6.72e-12). In total there were 14 OTUs that were common
to both models and the majority of these classified to commensal
bacteria (Bacteroides, Blautia, Streptococcus, and Clostridiales).\\
\textbf{Conclusions:} Our data suggests that treatment not only
significantly reduces the probability of having a colonic lesion within
our models but also causes detectable changes in the bacterial
microbiome. Further surveillance of these individuals will enable us to
determine whether models such as the ones we present could be used to
predict recurrence of colorectal cancer.
